---
title: "Day 1 - R and git intro"
author: "Casey Bowden"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Murray's stats 2023

<https://pcinereus.github.io/SUYRs_docs/>

## R studio

console for R

terminal for system

### install packages in Rgui

don't install through r studio

-   packages have dependencies which r studio holds in memory

-   r cant update what is in memory

install in terminal with R - pick first option

open Rgui and install in terminal - cant get it to work for me

#### check if you have a package

in r studio

```{r}
library(tidyverse)
library(car)
```

### shortcut keys

alt - shift - k to get shortcut keys

## polymorphism

runs a derivative of the function based on input you provide

ie mean; summary

```{r}
mean(x = 1:5)
```

## Git: version control

stored copy of all code so it can be reproduced

good for collaboration or auditing by reviewers

Git is a platform that allows us to keep track of versions of code and colaborate

only keeps track of changes

versions are called "**snapshots**" or "**commits**" and we can go back to any of these states at any time

the **head commit** is the one that will be build upon in future changes

#### System consists of three "trees"

-   **workspace** - files you see on your computer

-   **staging area** - where you nominate which files will be included in the next snapshot

    -   not every file you want to be included in version control

-   **local repository** - where you commit files too - but you cant see this tree

#### file can be in 4 states

-   **untracked** - unsafe, if you delete this file it is gone

-   **staged**- not safe yet but its on its way

-   **committed** - you could delete the file and get it back because it is committed to the snapshot

-   **modified** - when you change committed files - system knows you are keeping track of this file but it has been changed

#### see if git is installed

in terminal quit r q() and type git

in a new machine you need to tell git who you are

I can skip this step this time

in terminal type

git config --global user.name "Your name"

git config --global user.email "Your email"

check what your name is :

\$ git config \--global \--list

user.name=Casey Bowden

user.email=casey.bowden\@my.jcu.edu.au

Tell git to call repository main not master so that github knows where to look

git config \--global init.defaultBranch main

**to get git - click box when opening new project**

### .gitignore

any files listed in this file will not be commited without a big effrot

click on it to see

### git tab

can see status of various files

dont ant to commit project file but it will always appear on list as untracked

so add it to git ignore

### add project file to .gitignore

type into .gitignore

click save

it will disappear from git tab

### folder with data

dont want under version control or it to go to github

can put a folder in gitignore

ie if i wrote data/ anything in the folder called data will be ignored

### commit gitignore

1.  stage it (click the tick box)
    1.  changes to A for "added"

    2.  can click box again to unstage it
2.  commit
    1.  click commit button

    2.  first commit call "initial commit" in message

        1.  don't go looking for earlier commits, they don't exist
3.  box pops up to confirm commit

#### check commit history

click history within commit window

SHA is identification for the commit - often how you refer to it

-   only shows you first 7 character but it is 256 characters long

-   but often refer to by just first numbers

-   can see more of it underneath

##### create file to commit (ie commit this file)

```{r}
x=seq(1, 10, len=1)
y=40*2 + rnorm(10,0,5)
plot(x,y)

```

**COMMIT AND COMMIT OFTEN**

#### Split long script into smaller chunks

```{r}
mean(x)

```
